---
title: "homework 3, atomic vectors"
format: html
---

-----------------------------------------
Sof Antelo's CompBio Lab Homework 3 
1/28/26
Atomic Vectors Lab
-----------------------------------------

PART ONE: assigning values to the variables x, a, and b
```{r}
x = 1.1
a = 2.2
b = 3.3
```
-1.a
```{r}
z <- x^a^b 
print (z) 
```
object z assigned the value of x raised to a raised to b

-1.b
```{r}
z <- (x^a)^b 
print(z)
```
object z assigned the value of (x raised to a) raised to b

-1.c
```{r}
z <- (3*x^3) + (2*x^2) + 1 
print(z) 
```
assigning the equation 3x^3 + 2x^2 + 1 , to z

PART TWO
-2.a
```{r}
seq1 <- seq(from = 1, to = 8, length = 8)
seq2 <- seq(from = 7, to = 1, length = 7)
vec1 <- c(seq1, seq2)
print(vec1) #1 2 3 4 5 6 7 8 7 6 5 4 3 2 1
```
made two sequences, one forward 1-8, one backwards 7-1, then combined them into one vector using c()

-2.b
```{r}
vec2 <- rep (1:5, 1:5) 
print(vec2) 
```
makes a vector that goes from 1 to 5 [first 1:5], where the value repeats by the number of its place in vector [second 1:5, first shows 1, second twice, third thrice, etc.]

-2.c
```{r}
vec3 <- rep(5:1, 1:5)
print(vec3) 
```
by changing the first one from 1:5 to 5:1 we make 5 our starting number and 1 the end number, without changing the nature of the repition

PART THREE

-random vector of 2 nummbers, used set,seed so it is consistent
```{r}
set.seed(100)
rndm <- runif(2)
print(rndm) 
```

pythagoreon theory
```{r}
r <- sqrt(rndm[1]^2+rndm[2]^2) 
print(r) 
theta <- atan(rndm[2]/rndm[1]) 
print(theta) 
```

-polar coordinates
```{r}
polarCoordinates <- c(r, theta) 
print(polarCoordinates) 
```
vector of r, and theta 


PART FOUR: Noah's Ark

```{r}
queue <- c("sheep", "fox", "owl", "ant") 
print(queue)  
```
animals lined up to go on noahs ark sheep first, ant last

-4.a serpent gets in line
```{r}
queue[5] <- "serpent" 
print(queue) 
```
added a 5th posistion to queue and assigned it "serpent"

-4.b sheep enters the arc
```{r}
queue <- queue[-1]
print(queue) 
```

-4.c Donkey uses charm speak to cut the line 
```{r}
queue <- c("donkey", queue) 
print(queue) 
```
concatenates "donkey" and queue (makes vector with serpent in frist position and then the queue object lists everything else)

-4.d serpent has places to be
```{r}
queue <- queue[-5]
print(queue) 
```

-4.e owl has adhd and dips 
```{r}
queue <- queue[-3]
print(queue) 
```

4.f ant lets aphid cut 
```{r}
queue <- append(queue, 'aphid', after=2) 
print(queue)   
```
appending queue to put aphid after place 2

-4.g what place is aphid in? (without counting)
```{r}
which(queue == 'aphid') 
```

PART FIVE:create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.
```{r}
vec5 <- seq(1:100) 
vec5 <- vec5[!(vec5 %%2 == 0 | vec5 %%3 == 0 | vec5 %%7 == 0)] 

print(vec5) 
```
in order to make sure it was numbers NOT divisible have to put the parameter in () and have ! ahead of that within the [] so the ! will apply to each term separated by | for or
