---
title: "Homework 13"
format: html
---

-----------------------------------------
Sof Antelo's CompBio Lab Homework 13
4/15/26
GLMMs
-----------------------------------------

ASSIGNMENT: Using a dataset of your choice, ask two separate questions about the data and construct an appropriate model and figure to test and visualize each one.

For both questions I am going to use the bee dataset from the GLMMs lecture

Question 1: Can Site be a predictor for Nosema Load?
  Hypothesis: Nosema is transmitted through visiting fowers; there should be a relationship between nosema load and site.
  Null Hypothesis: There is not a relationship between site and nosema load.

```{r}
# load packages
library(tidyverse)
library(lubridate)
library(lme4)
library(car)

#loading the data
bee_data <- read_csv("data/Burnham_field_data_pathogens_wide.csv")
bee_dat <- bee_data %>%
  mutate(
    sampling_date = mdy(sampling_date),
    site_code = factor(site_code),
    field_id = factor(field_id),
    bee_caste = factor(bee_caste),
    bombus_spp = factor(bombus_spp),
    host_plant = factor(host_plant),
    sampling_event = factor(sampling_event),
    sampling_event_num = as.numeric(as.character(sampling_event)),
    log10_BQCV_load = log10(BQCV_pathogen_load + 1),
    log10_DWV_load = log10(DWV_pathogen_load + 1),
    log10_Nosema_load = log10(Nosema_pathogen_load + 1)
  )

glimpse(bee_dat)
df_filtered <- bee_dat[bee_dat$log10_Nosema_load > 0 & bee_dat$log10_BQCV_load > 0, ]

Nosema_by_field <- glm(
  log10_Nosema_load ~ field_id,
  data = df_filtered)

by_field <- Anova(Nosema_by_field)
print(by_field)


qplot(
  x = field_id,
  y = log10_Nosema_load,
  data = df_filtered) +
geom_boxplot(method = "lm", se = TRUE)

```
p-value of 0.03056 is statictically significant, meaning we can reject the null hypothesis and accept the alternative hypothesis.

Question 2: Can species be a predictor for BQCV?
  Hypothesis: Species is a predictor for BQCV.
  Null: Species does not impact BQCV load.
```{r}
bqcv_site <- lmer(
  log10_BQCV_load ~ bombus_spp + (1 | sampling_event),
  data = df_filtered)
Anova(bqcv_site)

#(1 | sampling_event) takes random sampling effects into account
df_filtered <- bee_dat[bee_dat$log10_Nosema_load > 0 & bee_dat$log10_BQCV_load > 0, ]


g2 <- ggplot(data=df_filtered) +
          aes(x=log10_BQCV_load) +
          geom_bar(fill="tomato",color="black")+ 
          theme(legend.position="none")
  theme(legend.position = "top")
  theme(legend.position = c(.2, .8))
  print(g2)   
```
p-value of 0.9161 is not statictically significant, meaning we fail to reject the null hypothesis and reject the alternative hypothesis. This means we cannot consider species to be a predictor for bqcv load. 