---
title: "Homework 4"
format: html
---
-----------------------------------
Sof Antelo's ComBio Lab Homework 4
2/4/26
-----------------------------------

Working with matrices, lists, and data frames

Part 1:
-Assign variable n_dims a single random integer 3:10.
```{r}
n_dims <- sample(3:10,1)
print(n_dims) 
```

-Create a vector of consecutive integers from 1 to n_dims2
```{r}
n_dims_vec <- c(1:(n_dims^2))
print(n_dims_vec) 
```

-Use sample function to randomly reshuffle
```{r}
n_dims_vec2 <- sample(n_dims_vec)
print(n_dims_vec2) 
```

-Create a square matrix with these elements.

```{r}
my_matrix <- matrix(data = n_dims_vec2, nrow = n_dims, ncol = n_dims )
print(my_matrix) 
```

-Find a function in r to transpose the matrix, print, note changes
```{r}
my_matrix2 <- t(my_matrix)

print(my_matrix2)
```

It seems like the rows and columns have switched i.e. what was column 1 is now row 1

-Calculate sum + mean of elements in the first + last row, with transposed matrix
```{r}
sum(my_matrix2[1,])  
sum(my_matrix2[n_dims,])
mean(my_matrix2[1,])
mean(my_matrix2[n_dims,]) 
```

-Exploring the eigen function
```{r}
ev <- eigen(my_matrix2)
print(ev)
```

-Elements of $values and $vectors
```{r}
typeof(ev$values)
typeof(ev$vectors)
```

Part 2: 
Create a list with the following named elements
-my_matrix, which is a 4 x 4 matrix filled with random uniform values
```{r}
my_matrix <- matrix(data=runif(16), nrow=4, ncol=4)
print(my_matrix)
```

-my_logical which is a 100-element vector of TRUE or FALSE values
```{r}
v <- runif(100)
my_logical <- v>.67
print(my_logical)
```

-my_letters, which is a 26-element vector of all the lower-case letters in random order
```{r}
my_letters <- c(sample(letters, size = 26, replace = FALSE))
print(my_letters)
```

*Subsequent Steps* 

-create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
my_list <- list(my_matrix[2,2],
                my_logical[2],
                my_letters[2])
print(my_list)
```

-Use the typeof() function to confirm the underlying data types of each component in this list
combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
typeof(my_list[[1]])
typeof(my_list[[2]])
typeof(my_list[[3]])
```

-Combine elements into a single atomic vector with c()
```{r}
my_vector <- c(my_list[[1]], my_list[[2]], my_list[[3]])
print(my_vector)
```

-What data typ is this vector?
```{r}
typeof(my_vector)
```

Part 3: Create a data frame with the two variables (= columns) and 26 cases (= rows)

```{r}
my_unis <- runif(26, min = 0, max = 10)
my_letters <- c(sample(LETTERS, size = 26, replace = FALSE))
my_df <- data.frame(my_unis, my_letters, stringsAsFactors = FALSE)
head(my_df)
```

-For the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA
```{r}
my_df[sample(length(my_unis), size = 4, replace = FALSE),1] <- NA
```

-For the first variable, write a single line of R code to identify which rows have the missing values.
```{r}
which(!complete.cases(my_df$my_unis))
```

-Re-order the entire data frame to arrange the second variable in alphabetical order
```{r}
my_df <- my_df[order(my_df$my_letters),]
head(my_df)
```

-Calculate the column mean for the first variable.
```{r}
mean(my_df$my_unis, na.rm = TRUE)
```