---
title: "Homework 8"
format: html
---

-----------------------------------------
Sof Antelo's CompBio Lab Homework 8
3/04/26
dplyr
-----------------------------------------

For loops and randomization tests

Set up:
```{r}
library(ggplot2)
library(tidyverse)
set.seed(1000)
```

PART 1: Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.
```{r}
zeroes <- function(
    myvec = sample(x= rep(0:4, times = 10), size = 10))
    {
        counter <- 0
        for(i in 1: length(myvec)){
            num <- myvec[i]
            ifelse(
                test = num == 0,
                yes = counter <-  counter + 1,
                no = counter <- counter
            )
        }
        return(counter)
}
set.seed(500)
zeroes()
```

PART 2: Use subsetting instead of a loop to rewrite the function as a single line of code.
```{r}
library(dplyr)
zero2 <- function(myvec = sample(x = rep(0:4, times = 10), size = 10)){
    results <- subset(myvec, myvec == 0)
    return(length(results))
}
print(zero2())
```

PART3: Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.
```{r}
mymat <- function(
    row = 12,
    col = 12
) {
    vec_row <- 1:row
    vec_col <- 1:col
    mat <- matrix(
        nrow = row,
        ncol = col
    )
    for(i in 1:row){
        for(j in 1:col){
            mat[i,j] <- vec_row[i] * vec_col[j]
        }
    }
    return(mat)
}
mymat()
```
PART 4: In the next few lectures, you will learn how to do a randomization test on your data. We will complete some of the steps today to practice calling custom functions within a for loop. Use the code from the March 31st lecture (Randomization Tests) to complete the following steps:

a. Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.
```{r}
set.seed(100)
groups <- rep(c("A", "B", "C"), each = 10)
response <- c(
    rnorm(n = 10, mean = 1),
    rnorm(n = 10, mean = 7),
    rnorm(n = 10, mean = 10)
)
df <- data.frame(
    group = groups,
    response = response
)
```

b. Write a custom function that 1: reshuffles the response variable, and 2: calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.
```{r}
library(dplyr)
shuffler <- function(df = df){
    reshuffled <- sample(df$response, size = nrow(df))
    df2 <- data.frame(groups = df$group, response = reshuffled)
    summary <- df2 |>
        group_(groups) |>
        summarize(meanResponse = mean(response))
    means_save <- summary <- summary$meanResponse
    return(means_save)
}
shuffler()
```

c. Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.
```{r}

```

d. Use qplot() to create a histogram of the means for each reshuffled group. Or, if you want a challenge, use ggplot() to overlay all 3 histograms in the same figure. How do the distributions of reshuffled means compare to the original means?
```{r}

```